### 2014-03-07 docs/related-work.tex
Fill some paragraphs and slightly changed the structure.

\documentclass{article} 
\usepackage{url}

\begin{document}

% VPN arms race We seek to build a system that remains difficult to block even
% after it has many users.

\section{Related work}

In response to Internet censorship, many research papers and pragmatic systems
such Freegate, Psiphon, and Lantern have explored the design space of censorship
circumvention. All these schemes are based on a simple idea: deploy a set of
proxies outside the censor's domain, and let the user connect to one of them
which can pass data on to the true destination for the user.

Broadly speaking, there are two main challenges to proxy-based circumvention.
The first is obfuscation: making circumvention traffic difficult to distinguish
from traffic the censor wishes to allow. The second is proxy address
unblockability: making it difficult for the censor simply to block proxies by
address. Past works have considered one or both of these problems.

To protect the relays from being detected and blocked, a school of research
tries to use covert channels for the communication with relays. FreeWave~\cite{freewave}
encodes data into audio streams and use the Skype P2P network to communicate
with the proxy located outside the censored domain, which is essentially another
Skype client that decodes the stream. The proxy then forward the traffic to the
true destination. A technique known as OSS~\cite{oss}, for ``online scanning
systems,'' closely resembles ours in important ways. It uses a third-party web
service as a conduit to carry traffic, encoded either in a URL or in an HTTP
body. Because web services cannot in general be assumed to reproduce a web
server's response exactly (a translation service, for example, returns a
response after translating it to another language), OSS does not use response
bodies to carry data, instead requiring the relay to initiate a reflected
request back to the client. OSS also does not claim that any one service is
difficult to block, rather that there are many web services that can be used.
The main ways in which this work differs is that we effectively control the OSS:
we can count on response bodies being preserved and so use them to send
downstream data; and we put the OSS on infrastructure that is expensive to
block.

Traffic using covert channels may exhibit distinguishable characteristics. As
noted by Winter et al.~\cite{foci12-winter}, Tor traffic is currently
distinguishable from what is regarded as harmless traffic. Therefore, another
school of research focuses on obfuscation techniques for hiding protocol
patterns. One technique is to look unlike anything forbidden by the censor; that
is, fail to match a blacklist. Another is to resemble a protocol that is
explicitly allowed; that is, match a whitelist. Falling into the first category
are ``look-like-nothing'' transports whose payloads are indistinguishable from a
uniformly random byte stream. A examples of such a transport is
obfsproxy~\cite{obfsproxy}, long the go-to obfuscated transport used by Tor.
ScrambleSuit~\cite{scramblesuit} is like obfsproxy in the content of its
payloads, but takes additional steps to obscure its signature of packet lengths
and timing, and is designed to resist active scanning for proxies (the proxy
server says nothing until the client proves knowledge of a shared secret).
Another path to obfsucation is the steganographic approach: look like something
the censor doesn't block. StegoTorus~\cite{stegotorus} and SkypeMorph~\cite{skypemorph} camouflage Tor
traffic and attempt to mimic HTTP or Skype traffic.

Houmansadr et al.~\cite{parrot} evaluate these steganography-based (parrot)
systems and conclude that unobservability by imitation is a ``fundamentally
flawed approach'', since these kind of systems must successfully mimic another
protocol. To fully mimic a complex and sometimes proprietary protocol like Skype
is difficult in that the system must mimic the protocol,  reaction to errors and
network conditions, typical traffic, and implementation-specific artifacts
correctly, otherwise a weak censor can recognize such traffic at a low cost.
Therefore we avoid the parrot approach and use HTTPS as the channel. Geddes  et
al.~\cite{acks} demonstrate that even non-parrot systems may be vulnerable to
attacks that disrupt that covert communication while having no effect or less
effect on legitimate traffic. Their examination is specific to VoIP protocols,
where packet losses and duplication are acceptable. The censor then can
deliberately drop or inject ACKs to disrupt the covert channel without causing
much collateral damage.

Tor has long faced the problem of its entry relays being blocked. The list of
relays is public, so it easy to block all of them by IP address. Tor
bridges~\cite{tor-blocking} are relays that are not universally known, intended
to serve as entry points for censored users. A system called BridgeDB seeks to
provide a few bridges to anyone who asks, while at the same time making it
difficult to learn the entire list of secret bridges. The plain Tor protocol
remains relatively easy to detect, so BridgeDB is also capable of distributing
obfsproxy and ScrambleSuit bridges.

telex censorspoofer no direction home

Flash proxy~\cite{flashproxy-pets12} attempts to address the problem of proxy
diversity by conscripting web users as temporary proxies. Proxies last only as
long as a web user stays on a page, so the pool of proxies is constantly
changing and difficult to block. Their approach to proxy unblockability is in a
sense opposite to this work's: where flash proxy uses many, cheap, unreliable
proxies, we use just one high-value proxy, to block which would cause expensive
collateral damage. There exists a prototype transport that combines flash proxy
with obfsproxy~\cite{obfs-flash}, however it is limited because it is not
possible to obfuscate flash proxy's outermost WebSocket layer.

How the GFW is blocking Tor. \cite{foci12-winter}

GoAgent

% CORDON? Say how meek would be classified in it?

\bibliographystyle{plain} bibliography{related-work}

\end{document}